High-Performance Language Primitives for Distributed Architectures using Asynchronous Events

### 2012-07-23 doc/asplos2013/interface.tex
2.4 and 2.5

\begin{lstlisting}[float={t},label={lst:regionapi},caption={Physical Region Interface and Example.}]
class RegionMetaData {
public:
  const unsigned id;
  static const RegionMetaData NO_REGION;
  // Create and destroy metadatas
  static RegionMetaData create_region(size_t num_elmts, 
                                      size_t elmt_size);
  void destroy_region() const;
  // Create and destroy instances
  RegionInstance create_instance(Memory memory) const;
  void destroy_instance(RegionInstance instance) const;
  // Create a physical instance only for reductions
  RegionInstance create_instance(Memory memory, 
                        ReductionOpID redopid) const;
};

@@

Listing~\ref{lst:regionapi} shows a subset of the interface for physical regions.  The
first object in the interface is a {\tt RegionMetaData} object (line 1).  Physical
regions created by the same region meta-data object belong to the same class.  Physical
regions are represented by {\tt RegionInstance} objects (line 17).  When creating a physical
region the client must specify the {\tt Memory} in which the physical region is going
to be allocated.  Memories are described in more detail in Section~\ref{subsec:machmodel}.
If there is not enough space for the physical region allocation a {\tt NO\_INST} will
be returned.  There is no automatic coherence of data between physical regions in the
same class.  It is the client's responsibility to manage data coherence using copies
between physical regions (line 22-23).  Just like other operations copies can be
predicated on an event and return an event corresponding to completion.

If reductions are the only operations that will be performed on locations in a region instance,
a special reduction version can be created by providing the ID of a {\em reduction operator}
when the object is created (line 13).  Reduction-only region instances can 
increase the available parallelism and hide latency better, as we will show in 
section~\ref{subsec:reducmicro}.  The interface supports any reduction operation that
is commutative (i.e. multiple operations to the same location can be reordered without 
changing the final result) and is coded as an {\tt apply} method as shown on line 29.
Many commonly-used reduction operations are also {\em foldable} (e.g. $(l \text{ += } r_1) \text{ += } r_2$ can
be folded as $l \text{ += } r_1 + r_2$), and if the {\tt fold} operation and a {\em right identity}
(e.g. $0$ is a right identity for addition because $x + 0 = x$ for all $x$) are provided,
the runtime can use a {\em reduction fold instance}, described in section~\ref{subsec:reducimpl},
for ever better performance.  Although the way they store data is very different,
reduction-only instances can be copied to other reduction-only instances (or normal 
instances) of the same class with the same asynchronous copy operations.  Copying between
instances with different layouts is only possible because physical regions provide 
the necessary information to understand the structure of the data.

@@

An example using reduction-only regions is shown on lines 35-49 of Listing~\ref{lst:regionapi}.
Two reduction-only regions, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 40-41).  A physical region of the same class {\tt inst3}
is also created in a different memory (line 42).  Two tasks are launched that will apply
reductions locally into their respective reduction buffers (lines 43-44).  The resulting
events are then used to chain copies back to {\tt inst3} after the tasks complete (45-46).  As
part of these copies the reductions stored in {\tt inst1} and {\tt inst2} will be applied
to {\tt inst3}.  Since reductions are done atomically, there is no race between the reduction
applications.  Finally, a merge on the events from the movement and reduction applications is 
performed and a final task using {\tt inst3} is launched (lines 47-48).  The resulting event
graph is shown in Figure~\ref{fig:reducevents}.

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is shown in Listing~\ref{lst:machineapi} and allows the
initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (line 15-16).
Any running task can get a pointer to the {\tt Machine} object with the
{\tt get\_machine} method (line 18) and use it to obtain lists of all the {\tt Memory} (line 19) or 
{\tt Processor} (line 20) handles in the system.  Using the affinity methods (lines 22-25),
the application can also determine which {\tt Memory} objects are accessible by each
{\tt Processor} and with what performance, and which pairs of {\tt Memory} objects are able to
efficiently copy data between them.

@@

\begin{lstlisting}[float={t},label={lst:machineapi},caption={Machine Interface.}]
class Memory {
public:
  const unsigned id;
  size_t size() const;
};

class Machine {
public:
  Machine(int *argc, char ***argv,
          const Processor::TaskIDTable &task_table);
  enum RunStyle {
    ONE_TASK_ONLY,ONE_TASK_PER_NODE,
    ONE_TASK_PER_PROCESSOR,
  };
  void run(Processor::TaskFuncID task_id = 0, 
        RunStyle = ONE_TASK_ONLY, const void *args, size_t arglen);
public:
  static Machine* get_machine(void);
  const set<Memory>& get_all_memories(void) const;
  const set<Processor>& get_all_processors(void) const;

>>>>

\begin{lstlisting}[float={t},label={lst:regionapi},caption={Physical Region Interface and Example.}]
class Memory {
public:
  const unsigned id;
  size_t size() const;
};

class RegionMetaData {
public:
  const unsigned id;
  static const RegionMetaData NO_REGION;
  // Create and destroy metadatas
  static RegionMetaData create_region(size_t num_elmts, 
                                      size_t elmt_size);
  void destroy_region() const;
  // Create and destroy instances
  RegionInstance create_instance(Memory memory) const;
  void destroy_instance(RegionInstance instance) const;
  // Create a physical instance only for reductions
  RegionInstance create_instance(Memory memory, 
                        ReductionOpID redopid) const;
};

@@

Listing~\ref{lst:regionapi} shows a subset of the interface for physical regions.  
Each memory in the machine's memory hierarchy is named by a {\tt Memory} object (line 1).
The next object is a {\tt RegionMetaData} object (line 7), which defines a class of
regions.  Physical regions that belong to that class are created with the {\tt create\_instance}
method (line 16) and represented by {\tt RegionInstance} objects (line 23).  When creating a physical
region the client specifies the memory in which the physical region is going
to be allocated.  To maintain performance transparency, a new physical region can only be allocated
if the memory has sufficient capacity.
There is no automatic coherence of data between physical regions in the
same class.  It is the client's responsibility to manage data coherence using copies
between physical regions (line 28-29).  Like other operations copies can be
predicated on an event and return an event corresponding to completion.

If reductions are the only operations that will be performed on locations in a region instance,
a special reduction version can be created by providing the ID of a {\em reduction operator}
when the object is created (line 19).  Reduction-only region instances can 
increase the available parallelism and hide latency better, as we will show in 
section~\ref{subsec:reducmicro}.  The interface supports any reduction operation that
is commutative (i.e. multiple operations to the same location can be reordered without 
changing the final result) and is coded as an {\tt apply} method as shown on line 35.
Many commonly-used reduction operations are also {\em foldable} (e.g. $(l \text{ += } r_1) \text{ += } r_2$ can
be folded as $l \text{ += } r_1 + r_2$), and if the {\tt fold} operation and a {\em right identity}
(e.g. $0$ is a right identity for addition because $x + 0 = x$ for all $x$) are provided,
the runtime can use a {\em reduction fold instance}, described in section~\ref{subsec:reducimpl},
for ever better performance.  Although the way they store data is very different,
reduction-only instances can be copied to other reduction-only instances (or normal 
instances) of the same class with the same asynchronous copy operations.  Copying between
instances with different layouts is only possible because physical regions provide 
the necessary information to understand the structure of the data.

@@

An example using reduction-only regions is shown on lines 41-55 of Listing~\ref{lst:regionapi}.
Two reduction-only regions, {\tt inst1} and {\tt inst2}, are created in different memories to
be used with a reduction operation (lines 46-47).  A physical region of the same class {\tt inst3}
is also created in a different memory (line 48).  Two tasks are launched that will apply
reductions locally into their respective reduction buffers (lines 49-50).  The resulting
events are then used to chain copies back to {\tt inst3} after the tasks complete (lines 51-52).  As
part of these copies the reductions stored in {\tt inst1} and {\tt inst2} will be applied
to {\tt inst3}.  Since reductions are done atomically, there is no race between the reduction
applications.  Finally, a merge on the events from the movement and reduction applications is 
performed and a final task using {\tt inst3} is launched (lines 53-54).  The resulting event
graph is shown in Figure~\ref{fig:reducevents}.

@@

\subsection{Machine}
\label{subsec:machmodel}
The last component of the interface is shown in Listing~\ref{lst:machineapi} and allows the
initialization and run-time introspection
of the current machine.
The {\tt Machine} object is a singleton object that the client creates at the start of a
program, providing a task table that maps task IDs to function pointers,
and then invoking the {\tt run} method (lines 9-10).
Any running task can get a pointer to the {\tt Machine} object with the
{\tt get\_machine} method (line 12) and use it to obtain lists of all the {\tt Memory} (line 13) or 
{\tt Processor} (line 14) handles in the system.  Using the affinity methods (lines 16-19),
the application can also determine which {\tt Memory} objects are accessible by each
{\tt Processor} and with what performance, and which pairs of {\tt Memory} objects are able to
efficiently copy data between them.

@@

\begin{lstlisting}[float={t},label={lst:machineapi},caption={Machine Interface.}]
class Machine {
public:
  Machine(int *argc, char ***argv,
          const Processor::TaskIDTable &task_table);
  enum RunStyle {
    ONE_TASK_ONLY,ONE_TASK_PER_NODE,
    ONE_TASK_PER_PROCESSOR,
  };
  void run(Processor::TaskFuncID task_id = 0, 
        RunStyle = ONE_TASK_ONLY, const void *args, size_t arglen);
public:
  static Machine* get_machine(void);
  const set<Memory>& get_all_memories(void) const;
  const set<Processor>& get_all_processors(void) const;
